---
title: "Rice Rivers Center - Impact Reporting"
date: "`r Sys.Date()`"
format: 
  html: 
    self-contained: true
    grid:
      margin-width: 350px
execute:
  echo: false
reference-location: margin
---

```{r}
#| output: false

# Load packages
library(googlesheets4)
library(dplyr)
library(tidyr)
library(lubridate)
library(strex)
library(knitr)
library(tidytext)
library(wordcloud)
```

```{r}
#| output: false

######################
### Authentication ###
######################

# Fetch the key from the system environment
gcp_key <- Sys.getenv("GCP_SA_KEY")

if (nzchar(gcp_key)) {
  # Authenticate using the github secret key
  gs4_auth(path = gcp_key)
  message("Authenticated using GCP Service Account Secret")
} else {
  # Authenticate using a browser if testing locally
  gs4_auth()
  message("Authenticated via browser.")
}

#################
### Read Data ###
#################

# Sheet ID for reservation request form responses:
sheet_id <- "1ed3zy3QERXUd7yMlfxtwxqAu3sqhpNWP-TScr1Y_dMM"
sheet_id_older_reservations <- "1nd2EMrJhVLAGLXVK840sgU7gEmqqZci1OvvG7t5Z5Hw"

reservations <- read_sheet(sheet_id)
older_reservations <- read_sheet(sheet_id_older_reservations)
```

```{r}

##################
### Processing ###
##################

# Subset reservations data to columns of interest
reservations <- reservations %>%
  select(`Requested arrival date`, `Requested departure date`, `Organization's Name`,
         `Description of Event`, `Facility Requested`, 
         `Will your group have VCU students attending?`,`How many VCU students will there be?`, 
         `Will your group have VCU employees attending?`,`How many VCU employees will there be?`,
         `Will your group have Non-VCU affiliated individuals attending?`,
         `How many Non-VCU affiliated individuals will there be?`, 
         `Is your group in the School of Life Sciences and Sustainability`, Status, 
         `Lodge Invoice Amount`, `Education Invoice Amount`, `Research Facility Invoice Amount`,
         `Fee Waived?`, `Waived Reason`)

# Combine old reservation data (pre-google form - based on invoices only) with new
# reservation data
reservations <- rbind(reservations, older_reservations)

# Formatting free text columns from list -> character
reservations$`How many Non-VCU affiliated individuals will there be?` <- as.character(reservations$`How many Non-VCU affiliated individuals will there be?`)
reservations$`How many VCU employees will there be?` <- as.character(reservations$`How many VCU employees will there be?`)
reservations$`How many VCU students will there be?` <- as.character(reservations$`How many VCU students will there be?`)


# Calculate additional columns and filter unfulfilled reservation requests:
reservations <- reservations %>%
  filter(Status == "Approved") %>%
  mutate(
    # Calculate number of days requested
    length_of_stay_days = as.numeric(difftime(`Requested departure date`, `Requested arrival date`, units = "days")) + 1,
    
    # Extract year
    year = year(`Requested arrival date`),
    
    # Approximate count of VCU students from free text response
    count_vcu_students_low = str_first_number(`How many VCU students will there be?`),
    count_vcu_students_high = str_last_number(`How many VCU students will there be?`),
    
    # Approximate count of VCU staff/faculty from free text response
    count_vcu_staff_low = str_first_number(`How many VCU employees will there be?`),
    count_vcu_staff_high = str_last_number(`How many VCU employees will there be?`),
    
    # Approximate count of non-VCU persons from free text response
    count_non_vcu_low = str_first_number(`How many Non-VCU affiliated individuals will there be?`),
    count_non_vcu_high = str_last_number(`How many Non-VCU affiliated individuals will there be?`)
    )

#########################
### Calculate Metrics ###
#########################

# 1. Count of students visiting per year

students_per_year <- reservations %>%
  group_by(year) %>%
  summarize(
    students_per_year_low_estimate = sum(count_vcu_students_low, na.rm = TRUE),
    students_per_year_high_estimate = sum(count_vcu_students_high, na.rm = TRUE)
  )

# 2. Count of VCU staff and faculty visiting per year

vcu_staff_per_year <- reservations %>%
  group_by(year) %>%
  summarize(
    vcu_staff_per_year_low_estimate = sum(count_vcu_staff_low, na.rm = TRUE),
    vcu_staff_per_year_high_estimate = sum(count_vcu_staff_high, na.rm = TRUE)
  )

# 3. Count of non-VCU visitors per year

non_vcu_per_year <- reservations %>%
  group_by(year) %>%
  summarize(
    non_vcu_per_year_low_estimate = sum(count_non_vcu_low, na.rm = TRUE),
    non_vcu_per_year_high_estimate = sum(count_non_vcu_high, na.rm = TRUE)
  )

# 4. Days, proportion, and estimated revenue per year of lodge usage

lodge_usage_days <- reservations %>%
  filter(grepl("Inger Rice Lodge", `Facility Requested`)) %>%
  group_by(year) %>%
  summarize(
    lodge_usage_days = sum(length_of_stay_days)
  )

lodge_usage_proportion <- lodge_usage_days %>%
  mutate(
    lodge_usage_proportion = as.numeric(lodge_usage_days)/365
  )

lodge_cost <- 15 # $15 per person per day

lodge_revenue <- reservations %>%
  filter(grepl("Inger Rice Lodge", `Facility Requested`)) %>% 
  filter(`Fee Waived?` == "Yes") %>%
  replace_na(list(count_vcu_students_low = 0, count_vcu_students_high = 0,
                  count_vcu_staff_low = 0, count_vcu_staff_high = 0,
                  count_non_vcu_low = 0, count_non_vcu_high = 0)
             ) %>%
  mutate(
    total_persons_low_estimate = as.numeric(count_vcu_students_low) + as.numeric(count_vcu_staff_low) + as.numeric(count_non_vcu_low),
    total_persons_high_estimate = as.numeric(count_vcu_students_high) + as.numeric(count_vcu_staff_high) + as.numeric(count_non_vcu_high)
  ) %>%
  mutate(
    total_persons_low_estimate = case_when(
      total_persons_low_estimate > 22 ~ 22,
      .default = total_persons_low_estimate
    ),
    total_persons_high_estimate = case_when(
      total_persons_high_estimate > 22 ~ 22,
      .default = total_persons_high_estimate
    )
  ) %>%
  mutate(
    revenue_per_event_low_estimate = length_of_stay_days*lodge_cost*total_persons_low_estimate,
    revenue_per_event_high_estimate = length_of_stay_days*lodge_cost*total_persons_high_estimate
  ) %>%
  group_by(year) %>%
  summarize(
    lodge_revenue_low_estimate = sum(revenue_per_event_low_estimate, na.rm = TRUE),
    lodge_revenue_high_estimate = sum(revenue_per_event_high_estimate, na.rm = TRUE)
    )

# 5. Days, proportion, and estimated revenue per year of conference room usage

conference_usage_days <- reservations %>%
  filter(grepl("Rice Education Building's Conference Room", `Facility Requested`)) %>%
  group_by(year) %>%
  summarize(
    conference_usage_days = sum(length_of_stay_days)
  )

conference_usage_proportion <- conference_usage_days %>%
  mutate(
    conference_usage_proportion = as.numeric(conference_usage_days)/365
  )

conference_cost <- 400

conference_revenue <- reservations %>%
  filter(grepl("Rice Education Building's Conference Room", `Facility Requested`)) %>% 
  filter(`Fee Waived?` == "Yes") %>%
  mutate(
    revenue_per_event = length_of_stay_days*conference_cost
  ) %>%
  group_by(year) %>%
  summarize(conference_revenue = sum(revenue_per_event))
  

# 6. Proportion of events with waved fees

      # Fees are waived for groups affiliated with SLSS

fees_waived <- reservations %>%
    filter(`Fee Waived?` == "Yes")

# 7. Paid Invoices

paid_reservations <- reservations %>%
  filter(!is.na(`Lodge Invoice Amount`) | !is.na(`Education Invoice Amount`) | !is.na(`Research Facility Invoice Amount`))

paid_invoices_by_year <- paid_reservations %>%
  group_by(year) %>%
  summarize(lodge_invoices = sum(`Lodge Invoice Amount`, na.rm=TRUE),
            education_invoices = sum(`Education Invoice Amount`, na.rm=TRUE),
            research_invoices = sum(`Research Facility Invoice Amount`, na.rm=TRUE))

```

This is a report of facility usage and engagement for the Rice Rivers Center. Data for this report
comes from the following sources:

1. Responses to the Rice Rivers Center [Reservation Form](https://ricerivers.vcu.edu/facilities/reservations/) (which handles reservations for RRC's
Inger Rice Lodge and Rice Education Building's Conference Room). 
2. Submissions to the Rice Rivers Center [Project Registration Form](https://ricerivers.vcu.edu/research/conduct-research-at-rice/project-registration/) (which handles research project
applications for VCU students, faculty, and staff.)

# Facilities Usage

```{r}

# Count of reservation requests by facility for the current year
current_year <- year(Sys.Date())

# Number of requests for the lodge
lodge_reservation_request_count <- as.numeric(nrow(reservations %>%
                                          filter(year == current_year & 
                                                   grepl("Inger Rice Lodge", 
                                                         `Facility Requested`))))

# Number of requests for the lodge
education_reservation_request_count <- as.numeric(nrow(reservations %>%
                                              filter(year == current_year & 
                                              grepl("Rice Education Building's Conference Room", 
                                                    `Facility Requested`))))

# Invoice total:
current_year_invoices <- paid_invoices_by_year %>%
  filter(year == current_year)

current_year_lodge_invoices <- current_year_invoices$lodge_invoices
current_year_lodge_invoices_formatted <- paste0("$", format(current_year_lodge_invoices, big.mark=","))
current_year_education_invoices <- current_year_invoices$education_invoices
current_year_education_invoices_formatted <- paste0("$", format(current_year_education_invoices, big.mark=","))

current_year_total_invoices <- current_year_lodge_invoices + current_year_education_invoices
current_year_total_invoices_formatted <- paste0("$", format(current_year_total_invoices, big.mark=","))

# Days booked
current_year_lodge_usage <- (lodge_usage_days %>%
  filter(year == current_year))$lodge_usage_days

current_year_education_usage <- (conference_usage_days %>%
  filter(year == current_year))$conference_usage_days

```

* In `{r} current_year`, there were `{r} lodge_reservation_request_count` reservation requests for the Inger Rice Lodge and `{r} education_reservation_request_count` reservation requests for the Education Building's Conference Room.
* The Inger Rice Lodge was booked for `{r} current_year_lodge_usage` days in `{r} current_year`. 
* The Education Building's Conference Room was booked for `{r} current_year_education_usage` days in `{r} current_year`.
* So far in `{r} current_year`, Rice Rivers Center facilities reservations have generated `{r} current_year_total_invoices_formatted` in revenue. 
  + Inger Rice Lodge reservations have generated `{r} current_year_lodge_invoices_formatted`.
  + Education Building's Conference Room reservations have generated `{r} current_year_education_invoices_formatted`.
  
Facilities have been utilized by VCU students, VCU staff & faculty, and non-VCU individuals. The table below provides an estimate on the number of individuals using Rice facilities each year. These estimates are based on responses from the the RRC Reservation Form. 

```{r}
# Merge tables 

usage_estimate <- merge(students_per_year, vcu_staff_per_year, by="year")
usage_estimate <- merge(usage_estimate, non_vcu_per_year, by="year")
usage_estimate <- usage_estimate %>%
  mutate(
    total_low_estimate = students_per_year_low_estimate + vcu_staff_per_year_low_estimate + non_vcu_per_year_low_estimate,
    total_high_estimate = students_per_year_high_estimate + vcu_staff_per_year_high_estimate + non_vcu_per_year_high_estimate
  ) %>%
  mutate(
    student_estimate = paste0(students_per_year_low_estimate, " - ", students_per_year_high_estimate),
    staff_estimate = paste0(vcu_staff_per_year_low_estimate, " - ", vcu_staff_per_year_high_estimate),
    non_vcu_estimate = paste0(non_vcu_per_year_low_estimate, " - ", non_vcu_per_year_high_estimate),
    total_estimate = paste0(total_low_estimate, " - ", total_high_estimate)
  ) %>%
  select(year, student_estimate, staff_estimate, non_vcu_estimate, total_estimate)

knitr::kable(usage_estimate,
             caption = "Count of individual usage of the Inger Rice Lodge and the Education Building Conference Room",
             format = "simple",
             col.names = c("Year", "VCU Students", "VCU Staff & Faculty", "Non-VCU Individuals", "Total"),
             align = c("l", "c", "c", "c", "r"))

```

```{r}

# Waived Usage Fees
current_year_waived_count <- nrow(fees_waived %>%
                                filter(year == current_year))
current_year_total_count <- nrow(reservations %>%
                                   filter(year == current_year))

```


Facility usage fees were waived for `{r} current_year_waived_count` out of `{r} current_year_total_count` total reservations in `{r} current_year`. Fees can be waived for non-profit groups, internal groups, or when there are issues with the facilities. Shown below is the breakdown of reasons for waived fees in `{r} current_year`. 

```{r}

waived_reason_proportion <- fees_waived %>%
  group_by(year, `Waived Reason`) %>%
  summarize(records = n()) %>%
  group_by(year) %>%
  mutate(proportion = paste0(format(round(records/sum(records)*100, 2), nsmall=2),"%")) %>%
  select(year, `Waived Reason`, proportion) %>%
  pivot_wider(
    names_from = `Waived Reason`, 
    values_from = proportion) %>%
  ungroup() %>%
  replace(is.na(.), "0%")

```




### VCU Student Engagement


```{r}
#| output: false

student_reservations <- reservations %>%
  filter(`Will your group have VCU students attending?` == "Yes")

# Proportion of student group reservations that use lodge, conference room, or both
student_reservation_facility_usage <- student_reservations %>%
  group_by(year, `Facility Requested`) %>%
  summarize(records = n()) %>%
  group_by(year) %>%
  mutate(proportion = paste0(format(round(records/sum(records)*100, 2), nsmall=2),"%")) %>%
  select(year, `Facility Requested`, proportion) %>%
  pivot_wider(
    names_from = `Facility Requested`, 
    values_from = proportion
  ) %>%
  ungroup() %>%
  replace(is.na(.), "0%") %>%
  rename(Both = `Rice Education Building's Conference Room, Inger Rice Lodge`)

```

```{r}

knitr::kable(student_reservation_facility_usage,
             caption = "Proportion of Facility Usage by Reservations incvolving VCU Students",
             format = "simple",
             col.names = c("Year", "Inger Rice Lodge Only", "Rice Education Building Only", "Both Facilities"),
             align = c("l", "c", "c", "c"))

```

\n
In 2026, the following groups reserved facilities at the Rice Rivers Center and brought VCU students:

```{r}

student_orgs <- unique(student_reservations$`Organization's Name`[student_reservations$year == 2026])
cat(paste("-", student_orgs), sep="\n")

```

### VCU Faculty & Staff Engagement

```{r}
#| output: false

staff_reservations <- reservations %>%
  filter(`Will your group have VCU employees attending?` == "Yes")

# Proportion of student group reservations that use lodge, conference room, or both
staff_reservation_facility_usage <- staff_reservations %>%
  group_by(year, `Facility Requested`) %>%
  summarize(records = n()) %>%
  group_by(year) %>%
  mutate(proportion = paste0(format(round(records/sum(records)*100, 2), nsmall=2),"%")) %>%
  select(year, `Facility Requested`, proportion) %>%
  pivot_wider(
    names_from = `Facility Requested`, 
    values_from = proportion
  ) %>%
  ungroup() %>%
  replace(is.na(.), "0%") %>%
  rename(Both = `Rice Education Building's Conference Room, Inger Rice Lodge`)

```

```{r}

knitr::kable(staff_reservation_facility_usage,
             caption = "Proportion of Facility Usage by Reservations involving VCU Faculty & Staff",
             format = "simple",
             col.names = c("Year", "Inger Rice Lodge Only", "Rice Education Building Only", "Both Facilities"),
             align = c("l", "c", "c", "c"))

```

\n
In 2026, the following groups reserved facilities at the Rice Rivers Center and brought VCU faculty and staff:

```{r}

staff_orgs <- unique(staff_reservations$`Organization's Name`[staff_reservations$year == 2026])
cat(paste("-", staff_orgs), sep="\n")

```
### Non-VCU Engagement

```{r}
#| output: false

non_vcu_reservations <- reservations %>%
  filter(`Will your group have Non-VCU affiliated individuals attending?` == "Yes")

# Proportion of student group reservations that use lodge, conference room, or both
non_vcu_reservation_facility_usage <- non_vcu_reservations %>%
  group_by(year, `Facility Requested`) %>%
  summarize(records = n()) %>%
  group_by(year) %>%
  mutate(proportion = paste0(format(round(records/sum(records)*100, 2), nsmall=2),"%")) %>%
  select(year, `Facility Requested`, proportion) %>%
  pivot_wider(
    names_from = `Facility Requested`, 
    values_from = proportion
  ) %>%
  ungroup() %>%
  replace(is.na(.), "0%") %>%
  rename(Both = `Rice Education Building's Conference Room, Inger Rice Lodge`)

```

```{r}

knitr::kable(non_vcu_reservation_facility_usage,
             caption = "Proportion of Facility Usage by Reservations involving Non-VCU Affiliated Individuals",
             format = "simple",
             col.names = c("Year", "Inger Rice Lodge Only", "Rice Education Building Only", "Both Facilities"),
             align = c("l", "c", "c", "c"))

```

\n
In 2026, the following groups reserved facilities at the Rice Rivers Center and brought individuals not affiliated with VCU:

```{r}

non_vcu_orgs <- unique(non_vcu_reservations$`Organization's Name`[non_vcu_reservations$year == 2026])
cat(paste("-", non_vcu_orgs), sep="\n")

```

##### Key Word Analysis of Visiting Groups

```{r}

tidytest <- reservations %>%
  select(`Description of Event`) %>%
  unnest_tokens(output = word, input = `Description of Event`)

numbers <- as.character(0:2030)

custom_stopwords <- tibble(
  word = c("vcu", "center", "space","fall", "rice", "building", "day", "first", "2026", "lodge", "need", "one", "rivers", 
           "aug", "available", "friday", "full", "know", "looking", "october", "others",
           "please", "probably", numbers),
  lexicon = "custom"
)

extended_stopwords <- stop_words %>%
  bind_rows(custom_stopwords)

cleaned_test <- tidytest %>%
  anti_join(extended_stopwords, by = "word")

cleaned_test %>%
  count(word, sort = TRUE) %>%
  with(wordcloud(word, n, max.words = 40))

```